---
title: Livestock Breeding and Genomics - Solution 8
author: "Peter von Rohr"
date: "2019-11-15"
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    keep_tex: false
header-includes:
 \usepackage{longtable}
 \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'asis', fig.pos = 'H')
knitr::knit_hooks$set(hook_convert_odg = rmddochelper::hook_convert_odg)
```

## Problem 1: Compute Inbreeding Coefficients
Given the following pedigree.

```{r}
n_nr_ani_ped <- 6
n_nr_parent <- 3
tbl_ped_sol10p01 <- tibble::tibble(Animal = c(1:n_nr_ani_ped),
                             Sire = c(NA, NA, 1, 3, 4, 4),
                             Dam  = c(NA, NA, NA, 2, 2, 5))
### # show the pedigree
knitr::kable( tbl_ped_sol10p01,
              booktabs = TRUE,
              longtable = TRUE )
```

### Your Task
Compute the inbreeding coefficients $F_i$ for all animals using the matrix $R$ that comes from the cholesky decomposition of the numerator relationship matrix $A$


### Solution
The cholesky decomposition of $A$ corresponds to 

$$A = R \cdot R^T$$

where $R$ is a lower triangular matrix. The diagonal elements $(A)_{ii}$ of $A$ can be computed as the sum of the squared elements of all elements of Matrix $R$ on row $i$. 

$$(A)_{ii} = \sum_{j=1}^i (R)_{ij}^2$$

Therefore to get $(A)_{ii}$, we have to compute all elements of $R$ on row $i$. Due to the relation of the cholesky-decomposition to the `LDL`-decomposition, we can write the matrix $R$ as a product of the two matrices $L$ and $S$

\begin{equation}
R = L \cdot S
(\#eq:decomprintols)
\end{equation}

where $L$ is the lower triangular matrix from the `LDL`-decomposition and $S$ is a diagonal matrix with diagonal elements $(S)_{ii}$ corresponding to 

$$(S)_{ii} = \sqrt{(D)_{ii}}$$

where $(D)_{ii}$ correspond to the diagonal elements of the matrix $D$ from the `LDL`-decomposition. From the `LDL`, we know that 

$$(D)_{ii} = {1\over 2} - {1\over 4}(F_s + F_d) = 1 - {1\over 4}(A_{ss} + A_{dd})$$

Based on the decomposition of $R$ into the product of $L$ and $S$ given in \@ref(eq:decomprintols) and based on the property of the matrix $L$ which is to be shown in the additional problem 3 of this exercise, we can derive the following rules for computing the elements of the matrix $R$

* Diagnoal elements $(R)_{ii}$:

$$(R)_{ii} = (S)_{ii} = \sqrt{(D)_{ii}} = \sqrt{1 - {1\over 4}(A_{ss} + A_{dd})}$$
where $s$ and $d$ are parents of animal $i$ and $(A)_{ss}$ and $(A)_{dd}$ are diagonal elements of the numerator relationship matrix $A$.

* Off-diagonal elements $(R)_{ij}$ ($i \ne j$):

$$(R)_{ij} = {1\over 2}(R_{sj} + R_{dj})$$

where $s$ and $d$ are parents of animal $i$. 

```{r compsolprep, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
suppressPackageStartupMessages( library(pedigreemm) )
ped_sol10p01 <- pedigree(sire = tbl_ped_sol10p01$Sire, dam = tbl_ped_sol10p01$Dam, label = as.character(tbl_ped_sol10p01$Animal))
matA_sol10p01 <- as.matrix(getA(ped = ped_sol10p01))
matR_sol10p01 <- t(chol(matA_sol10p01))
n_matA_na <- 0
n_matR_na <- 0
vec_Aii <- vector("numeric", nrow(tbl_ped_sol10p01))
```

The solution of this exercise is to compute $(A)_{ii}$ for all animals in the pedigree using the above described rules.

```{r compmatrfunc, echo=FALSE, results='hide'}
#' @title Get elements of a given row of matrix R
#'
get_mat_r_row_element <- function(pn_ani_idx, ptbl_ped){
  ### # get pedigree
  ped <- pedigreemm::pedigree(sire = ptbl_ped$Sire, 
                              dam = ptbl_ped$Dam, 
                              label = as.character(ptbl_ped$Animal))
  ### # compute matrix A from pedigree
  matA <- as.matrix(getA(ped = ped))
  matR <- t(chol(matA))
  n_matA_na <- 0
  n_matR_na <- 0
  ### # find sire and dam idx
  n_sire_idx <- ptbl_ped$Sire[pn_ani_idx]
  n_dam_idx <- ptbl_ped$Dam[pn_ani_idx]
  ### # diagonal element
  n_matA_sire <- n_matA_na
  if (!is.na(n_sire_idx))
    n_matA_sire <- matA[n_sire_idx, n_sire_idx]
  n_matA_dam <- n_matA_na
  if (!is.na(n_dam_idx))
    n_matA_dam <- matA[n_dam_idx, n_dam_idx]
  n_matRii2 <- 1 - .25 * (n_matA_sire + n_matA_dam)
  ### # check computed value of matRii with value from chol()
  if (class(all.equal(matR[pn_ani_idx, pn_ani_idx]^2, n_matRii2)) == "character")
    stop("[ERROR -- get_mat_r_row_element] Inconsistent diagonal element of R for animal: ", pn_ani_idx, "\n")
  ### # offdiagnoal element
  vec_matRij2 <- NULL
  ### # column index to run through on row pn_ani_idx
  vec_col_idx <- NULL
  if (n_ani_idx > 1)
    vec_col_idx <- c(1:(pn_ani_idx-1))
  for (colidx in vec_col_idx) {
    n_matR_sire <- n_matR_na
    if (!is.na(n_sire_idx))
      n_matR_sire <- matR[n_sire_idx, colidx]
    n_matR_dam <- n_matR_na
    if (!is.na(n_dam_idx))
      n_matR_dam <- matR[n_dam_idx, colidx]
    n_matR_ani <- 0.5 * (n_matR_sire + n_matR_dam)
    ### # check with real matrix R
    if (class(all.equal(matR[pn_ani_idx, colidx], n_matR_ani)) == "character")
      stop("[ERROR -- get_mat_r_row_element] Inconsitent off-diagnoal of R on row: ", pn_ani_idx, " and col: ", colidx, "\n")
    ### # aggregate results
    if (is.null(vec_matRij2)){
      vec_matRij2 <- n_matR_ani^2
    } else {
      vec_matRij2 <- c(vec_matRij2, n_matR_ani^2)
    }
  }
  ### # return list of results
  return(list(Rii2 = n_matRii2, Rij2 = vec_matRij2))
}
```


```{r compsola11, echo=FALSE, results='hide'}
n_ani_idx <- 1
l_matr_elem <- get_mat_r_row_element(pn_ani_idx = n_ani_idx, ptbl_ped = tbl_ped_sol10p01)
if (is.null(l_matr_elem$Rij2)){
  vec_Aii[n_ani_idx] <- l_matr_elem$Rii2
}else {
   vec_Aii[n_ani_idx] <- sum(l_matr_elem$Rij2) + l_matr_elem$Rii2
}
```

* $(A)_{`r n_ani_idx``r n_ani_idx`}$

$$(A)_{`r n_ani_idx``r n_ani_idx`} = (R)_{`r n_ani_idx``r n_ani_idx`}^2  = `r vec_Aii[n_ani_idx]`$$

```{r compsola22, echo=FALSE, results='hide'}
n_ani_idx <- 2
l_matr_elem <- get_mat_r_row_element(pn_ani_idx = n_ani_idx, ptbl_ped = tbl_ped_sol10p01)
if (is.null(l_matr_elem$Rij2)){
  vec_Aii[n_ani_idx] <- l_matr_elem$Rii2
}else {
   vec_Aii[n_ani_idx] <- sum(l_matr_elem$Rij2) + l_matr_elem$Rii2
}
```

* $(A)_{`r n_ani_idx``r n_ani_idx`}$

```{r showelema22, echo=FALSE, results='asis'}
cat("$$\n")
### # show A-element part
cat("(A)_{", n_ani_idx, n_ani_idx, "} = ", sep = "")
### # offdiagonals of R
if (!is.null(l_matr_elem$Rij2)){
  for (colidx in seq_along(l_matr_elem$Rij2)){
    cat("(R)_{", n_ani_idx, colidx, "}^2 + ", sep = "")
  }
}
cat("(R)_{", n_ani_idx, n_ani_idx, "}^2", sep = "")
cat(" = ")
if (!is.null(l_matr_elem$Rij2)){
  for (colidx in seq_along(l_matr_elem$Rij2)){
    cat(l_matr_elem$Rij2[colidx], " + ", sep = "")
  }
}
cat(l_matr_elem$Rii2)
cat(" = ")
cat(vec_Aii[n_ani_idx], "\n")
cat("$$\n")
```

```{r compsola33, echo=FALSE, results='hide'}
n_ani_idx <- 3
l_matr_elem <- get_mat_r_row_element(pn_ani_idx = n_ani_idx, ptbl_ped = tbl_ped_sol10p01)
if (is.null(l_matr_elem$Rij2)){
  vec_Aii[n_ani_idx] <- l_matr_elem$Rii2
}else {
   vec_Aii[n_ani_idx] <- sum(l_matr_elem$Rij2) + l_matr_elem$Rii2
}
```

* $(A)_{`r n_ani_idx``r n_ani_idx`}$

```{r showelema33, echo=FALSE, results='asis'}
cat("$$\n")
### # show A-element part
cat("(A)_{", n_ani_idx, n_ani_idx, "} = ", sep = "")
### # offdiagonals of R
if (!is.null(l_matr_elem$Rij2)){
  for (colidx in seq_along(l_matr_elem$Rij2)){
    cat("(R)_{", n_ani_idx, colidx, "}^2 + ", sep = "")
  }
}
cat("(R)_{", n_ani_idx, n_ani_idx, "}^2", sep = "")
cat(" = ")
if (!is.null(l_matr_elem$Rij2)){
  for (colidx in seq_along(l_matr_elem$Rij2)){
    cat(l_matr_elem$Rij2[colidx], " + ", sep = "")
  }
}
cat(l_matr_elem$Rii2)
cat(" = ")
cat(vec_Aii[n_ani_idx])
cat("$$\n")
```

```{r compsola44, echo=FALSE, results='hide'}
n_ani_idx <- 4
l_matr_elem <- get_mat_r_row_element(pn_ani_idx = n_ani_idx, ptbl_ped = tbl_ped_sol10p01)
if (is.null(l_matr_elem$Rij2)){
  vec_Aii[n_ani_idx] <- l_matr_elem$Rii2
}else {
   vec_Aii[n_ani_idx] <- sum(l_matr_elem$Rij2) + l_matr_elem$Rii2
}
```

* $(A)_{`r n_ani_idx``r n_ani_idx`}$

```{r showelema44, echo=FALSE, results='asis'}
cat("$$\n")
### # show A-element part
cat("(A)_{", n_ani_idx, n_ani_idx, "} = ", sep = "")
### # offdiagonals of R
if (!is.null(l_matr_elem$Rij2)){
  for (colidx in seq_along(l_matr_elem$Rij2)){
    cat("(R)_{", n_ani_idx, colidx, "}^2 + ", sep = "")
  }
}
cat("(R)_{", n_ani_idx, n_ani_idx, "}^2", sep = "")
cat("$$\n")
cat("$$\n")
cat(" = ")
if (!is.null(l_matr_elem$Rij2)){
  for (colidx in seq_along(l_matr_elem$Rij2)){
    cat(l_matr_elem$Rij2[colidx], " + ", sep = "")
  }
}
cat(l_matr_elem$Rii2)
cat(" = ")
cat(vec_Aii[n_ani_idx])
cat("$$\n")
```


```{r compsola55, echo=FALSE, results='hide'}
n_ani_idx <- 5
l_matr_elem <- get_mat_r_row_element(pn_ani_idx = n_ani_idx, ptbl_ped = tbl_ped_sol10p01)
if (is.null(l_matr_elem$Rij2)){
  vec_Aii[n_ani_idx] <- l_matr_elem$Rii2
}else {
   vec_Aii[n_ani_idx] <- sum(l_matr_elem$Rij2) + l_matr_elem$Rii2
}
```

* $(A)_{`r n_ani_idx``r n_ani_idx`}$

```{r showelema55, echo=FALSE, results='asis'}
cat("$$\n")
### # show A-element part
cat("(A)_{", n_ani_idx, n_ani_idx, "} = ", sep = "")
### # offdiagonals of R
if (!is.null(l_matr_elem$Rij2)){
  for (colidx in seq_along(l_matr_elem$Rij2)){
    cat("(R)_{", n_ani_idx, colidx, "}^2 + ", sep = "")
  }
}
cat("(R)_{", n_ani_idx, n_ani_idx, "}^2", sep = "")
cat("$$\n")
cat("$$\n")
cat(" = ")
if (!is.null(l_matr_elem$Rij2)){
  for (colidx in seq_along(l_matr_elem$Rij2)){
    cat(l_matr_elem$Rij2[colidx], " + ", sep = "")
  }
}
cat(l_matr_elem$Rii2)
cat(" = ")
cat(vec_Aii[n_ani_idx])
cat("$$\n")
```

```{r compsola66, echo=FALSE, results='hide'}
n_ani_idx <- 6
l_matr_elem <- get_mat_r_row_element(pn_ani_idx = n_ani_idx, ptbl_ped = tbl_ped_sol10p01)
if (is.null(l_matr_elem$Rij2)){
  vec_Aii[n_ani_idx] <- l_matr_elem$Rii2
}else {
   vec_Aii[n_ani_idx] <- sum(l_matr_elem$Rij2) + l_matr_elem$Rii2
}
```

* $(A)_{`r n_ani_idx``r n_ani_idx`}$

```{r showelema66, echo=FALSE, results='asis'}
cat("$$\n")
### # show A-element part
cat("(A)_{", n_ani_idx, n_ani_idx, "} = ", sep = "")
### # offdiagonals of R
if (!is.null(l_matr_elem$Rij2)){
  for (colidx in seq_along(l_matr_elem$Rij2)){
    cat("(R)_{", n_ani_idx, colidx, "}^2 + ", sep = "")
  }
}
cat("(R)_{", n_ani_idx, n_ani_idx, "}^2", sep = "")
cat("$$\n")
cat("$$\n")
cat(" = ")
if (!is.null(l_matr_elem$Rij2)){
  for (colidx in seq_along(l_matr_elem$Rij2)){
    cat(l_matr_elem$Rij2[colidx], " + ", sep = "")
  }
}
cat(l_matr_elem$Rii2)
cat(" = ")
cat(vec_Aii[n_ani_idx])
cat("$$\n")
```

As a check, we can compute the inbreeding coefficients using the function `pedigreemm::inbreeding()`

```{r echo=TRUE, results='markup'}
pedigreemm::inbreeding(ped = ped_sol10p01)
```


## Problem 2: Direct Construction of $A^{-1}$
Use the pedigree from problem 1 and the computed inbreeding coefficients from problem 1 to set up the inverse numerator relationship matrix $A^{-1}$ using the general form of Henderson's rules for a pedigree with inbred animals. Compare your result using function `pedigreemm::getAInv()`.


### Solution
As a pre-requisite, we assume that the pedigree is sorted such that parents come before progeny. Henderson's rules contain the following steps

* Start with a matrix $A^{-1}$ where all elements are set to $0$.
* Let $d^i$ be the $i$-th diagonal element of $D^{-1}$ for animal $i$, assuming $i$ has parents $s$ and $d$.
* Then add the following contributions to $A^{-1}$
    + $d^i$ to the element $(i,i)$
    + $-d^i/2$ to the elements $(s,i)$, $(i,s)$, $(d,i)$, $(i,d)$
    + $d^i/4$ to the elements $(s,s)$, $(s,d)$, $(d,s)$, $(d,d)$

Applying these rules to the pedigree given in problem 1 leads to the following sequence of computations.


* Initialize the matrix $A^{-1}$ with all $0$

```{r matainvinit}
n_nr_ani <- nrow(tbl_ped_sol10p01)
matAinv <- matrix(0, nrow = n_nr_ani, ncol = n_nr_ani)
cat("$$\n")
cat("A^{-1} = \\left[")
cat(paste(rmddochelper::sConvertMatrixToLaTexArray(pmatAMatrix = matAinv, pnDigits = 0)))
cat("\\right]\n")
cat("$$\n")
```


```{r hendersonrulfunctions}
#' @title Get diagonal element from D inverse for animal pn_ani_idx
get_diag_element_dinv <- function(pn_ani_idx, pn_sire_idx, pn_dam_idx, pvec_inbreeding){
  ### # pn_ani_idx cannot be NA
  if (is.na(pn_ani_idx))
    stop("[ERROR -- get_diag_element_dinv] Animal ID pn_ani_idx cannot be NA")
  ### # initialize inbreeding of parents to 0
  n_f_sire <- 0
  n_f_dam <- 0
  ### # if parents are known, take their inbreeding from pvec_inbreeding
  if (!is.na(pn_sire_idx)){
    n_f_sire <- pvec_inbreeding[pn_sire_idx]
  }
  ### # dam
  if (!is.na(pn_dam_idx)){
    n_f_dam <- pvec_inbreeding[pn_dam_idx]
  }
  
  ### # compute diagonal element of D
  if (!is.na(pn_sire_idx)){
    if (!is.na(pn_dam_idx)){
      n_diag_d <- 0.5 - 0.25*(n_f_sire + n_f_dam)
    } else {
      n_diag_d <- 0.75 - 0.25 * n_f_sire
    }
  } else {
    if (!is.na(pn_dam_idx)){
      n_diag_d <- 0.75 - 0.25 * n_f_dam
    } else {
      n_diag_d <- 1
    }
  }
  # return inverse of diagnoal element of D
  return(1/n_diag_d)
}

#' @title Update A inverse for a given animal
update_mat_ainv <- function(pn_ani_idx, pn_sire_idx, pn_dam_idx, pn_d, pmat_ainv){
  ### # copy the current matrix
  mat_inv_result <- pmat_ainv
  ### # pn_ani_idx cannot be NA
  if (is.na(pn_ani_idx))
    stop("[ERROR -- update_mat_ainv] Animal ID pn_ani_idx cannot be NA")
  ### # add pn_d to diagonal
  mat_inv_result[pn_ani_idx, pn_ani_idx] <- mat_inv_result[pn_ani_idx, pn_ani_idx] + pn_d
  ### # add contributions for sire, if sire is not NA
  if (!is.na(pn_sire_idx)){
    mat_inv_result[pn_ani_idx, pn_sire_idx] <- mat_inv_result[pn_ani_idx, pn_sire_idx] - pn_d/2
    mat_inv_result[pn_sire_idx, pn_ani_idx] <- mat_inv_result[pn_sire_idx, pn_ani_idx] - pn_d/2
    mat_inv_result[pn_sire_idx, pn_sire_idx] <- mat_inv_result[pn_sire_idx, pn_sire_idx] + pn_d/4
  }
  ### # add contributions for dam, if dam is not NA
  if (!is.na(pn_dam_idx)){
    mat_inv_result[pn_ani_idx, pn_dam_idx] <- mat_inv_result[pn_ani_idx, pn_dam_idx] - pn_d/2
    mat_inv_result[pn_dam_idx, pn_ani_idx] <- mat_inv_result[pn_dam_idx, pn_ani_idx] - pn_d/2
    mat_inv_result[pn_dam_idx, pn_dam_idx] <- mat_inv_result[pn_dam_idx, pn_dam_idx] + pn_d/4
    ### # both parents known, add off-diagonal contributions
    if (!is.na(pn_sire_idx)){
      mat_inv_result[pn_dam_idx, pn_sire_idx] <- mat_inv_result[pn_dam_idx, pn_sire_idx] + pn_d/4
      mat_inv_result[pn_sire_idx, pn_dam_idx] <- mat_inv_result[pn_sire_idx, pn_dam_idx] + pn_d/4
    }  
  }
  ### # return updated matrix
  return(mat_inv_result)
}
```

```{r ainvpreparation}
vec_inbreeding <- pedigreemm::inbreeding(ped = ped_sol10p01)
```

```{r animal1}
n_ani_idx <- 1
n_sire_idx <- tbl_ped_sol10p01$Sire[n_ani_idx]
n_dam_idx <- tbl_ped_sol10p01$Dam[n_ani_idx]
n_elem_dinv <- get_diag_element_dinv(pn_ani_idx = n_ani_idx, 
                                     pn_sire_idx = n_sire_idx, 
                                     pn_dam_idx = n_dam_idx, 
                                     pvec_inbreeding = vec_inbreeding)
matAinv <- update_mat_ainv(pn_ani_idx = n_ani_idx, 
                           pn_sire_idx = n_sire_idx,
                           pn_dam_idx = n_dam_idx,
                           pn_d = n_elem_dinv,
                           pmat_ainv = matAinv)
```

* Animal `r n_ani_idx`

```{r matainvanimal1}
cat("$$\n")
cat("A^{-1} = \\left[")
cat(paste(rmddochelper::sConvertMatrixToLaTexArray(pmatAMatrix = matAinv, pnDigits = 0)))
cat("\\right]\n")
cat("$$\n")
```

```{r animal2}
n_ani_idx <- 2
n_sire_idx <- tbl_ped_sol10p01$Sire[n_ani_idx]
n_dam_idx <- tbl_ped_sol10p01$Dam[n_ani_idx]
n_elem_dinv <- get_diag_element_dinv(pn_ani_idx = n_ani_idx, 
                                     pn_sire_idx = n_sire_idx, 
                                     pn_dam_idx = n_dam_idx, 
                                     pvec_inbreeding = vec_inbreeding)
matAinv <- update_mat_ainv(pn_ani_idx = n_ani_idx, 
                           pn_sire_idx = n_sire_idx,
                           pn_dam_idx = n_dam_idx,
                           pn_d = n_elem_dinv,
                           pmat_ainv = matAinv)
```

* Animal `r n_ani_idx`

```{r matainvanimal2}
cat("$$\n")
cat("A^{-1} = \\left[")
cat(paste(rmddochelper::sConvertMatrixToLaTexArray(pmatAMatrix = matAinv, pnDigits = 0)))
cat("\\right]\n")
cat("$$\n")
```

```{r animal3}
n_ani_idx <- 3
n_sire_idx <- tbl_ped_sol10p01$Sire[n_ani_idx]
n_dam_idx <- tbl_ped_sol10p01$Dam[n_ani_idx]
n_elem_dinv <- get_diag_element_dinv(pn_ani_idx = n_ani_idx, 
                                     pn_sire_idx = n_sire_idx, 
                                     pn_dam_idx = n_dam_idx, 
                                     pvec_inbreeding = vec_inbreeding)
matAinv <- update_mat_ainv(pn_ani_idx = n_ani_idx, 
                           pn_sire_idx = n_sire_idx,
                           pn_dam_idx = n_dam_idx,
                           pn_d = n_elem_dinv,
                           pmat_ainv = matAinv)
```

* Animal `r n_ani_idx`

```{r matainvanimal3}
cat("$$\n")
cat("A^{-1} = \\left[")
cat(paste(rmddochelper::sConvertMatrixToLaTexArray(pmatAMatrix = matAinv, pnDigits = 4)))
cat("\\right]\n")
cat("$$\n")
```

```{r animal4}
n_ani_idx <- 4
n_sire_idx <- tbl_ped_sol10p01$Sire[n_ani_idx]
n_dam_idx <- tbl_ped_sol10p01$Dam[n_ani_idx]
n_elem_dinv <- get_diag_element_dinv(pn_ani_idx = n_ani_idx, 
                                     pn_sire_idx = n_sire_idx, 
                                     pn_dam_idx = n_dam_idx, 
                                     pvec_inbreeding = vec_inbreeding)
matAinv <- update_mat_ainv(pn_ani_idx = n_ani_idx, 
                           pn_sire_idx = n_sire_idx,
                           pn_dam_idx = n_dam_idx,
                           pn_d = n_elem_dinv,
                           pmat_ainv = matAinv)
```

* Animal `r n_ani_idx`

```{r matainvanimal4}
cat("$$\n")
cat("A^{-1} = \\left[")
cat(paste(rmddochelper::sConvertMatrixToLaTexArray(pmatAMatrix = matAinv, pnDigits = 4)))
cat("\\right]\n")
cat("$$\n")
```

```{r animal5}
n_ani_idx <- 5
n_sire_idx <- tbl_ped_sol10p01$Sire[n_ani_idx]
n_dam_idx <- tbl_ped_sol10p01$Dam[n_ani_idx]
n_elem_dinv <- get_diag_element_dinv(pn_ani_idx = n_ani_idx, 
                                     pn_sire_idx = n_sire_idx, 
                                     pn_dam_idx = n_dam_idx, 
                                     pvec_inbreeding = vec_inbreeding)
matAinv <- update_mat_ainv(pn_ani_idx = n_ani_idx, 
                           pn_sire_idx = n_sire_idx,
                           pn_dam_idx = n_dam_idx,
                           pn_d = n_elem_dinv,
                           pmat_ainv = matAinv)
```

* Animal `r n_ani_idx`

```{r matainvanimal5}
cat("$$\n")
cat("A^{-1} = \\left[")
cat(paste(rmddochelper::sConvertMatrixToLaTexArray(pmatAMatrix = matAinv, pnDigits = 4)))
cat("\\right]\n")
cat("$$\n")
```

```{r animal6}
n_ani_idx <- 6
n_sire_idx <- tbl_ped_sol10p01$Sire[n_ani_idx]
n_dam_idx <- tbl_ped_sol10p01$Dam[n_ani_idx]
n_elem_dinv <- get_diag_element_dinv(pn_ani_idx = n_ani_idx, 
                                     pn_sire_idx = n_sire_idx, 
                                     pn_dam_idx = n_dam_idx, 
                                     pvec_inbreeding = vec_inbreeding)
matAinv <- update_mat_ainv(pn_ani_idx = n_ani_idx, 
                           pn_sire_idx = n_sire_idx,
                           pn_dam_idx = n_dam_idx,
                           pn_d = n_elem_dinv,
                           pmat_ainv = matAinv)
```

* Animal `r n_ani_idx`

```{r matainvanimal6}
cat("$$\n")
cat("A^{-1} = \\left[")
cat(paste(rmddochelper::sConvertMatrixToLaTexArray(pmatAMatrix = matAinv, pnDigits = 4)))
cat("\\right]\n")
cat("$$\n")
```


* Check with function `pedigreemm::getAinv()`

```{r matainvcheck, echo=TRUE, results='markup'}
pedigreemm::getAInv(ped = ped_sol10p01)
```

The difference between the computed matrix and the matrix from `pedigreemm::getAinv()`

```{r matainvdiff}
matAinv_check <- as.matrix(pedigreemm::getAInv(ped = ped_sol10p01))
cat("$$\n")
cat("\\left[")
cat(paste(rmddochelper::sConvertMatrixToLaTexArray(pmatAMatrix = matAinv_check-matAinv, pnDigits = 2)))
cat("\\right]\n")
cat("$$\n")
```

