---
title: Livestock Breeding and Genomics - Solution 4
author: Peter von Rohr
date: 2019-10-18
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'asis')
```

## Problem 1: Prediction of Breeding Value and Reliability
```{r SetUpEx1}
y <- 320
mu <- 250
h2 <- 0.45
```

Heifer Elsa has a weaning weight of $`r y`$ kg. The population mean of the same trait is at $`r mu`$ kg.The heritability ($h^2$) of the trait weaning weight is assumed to be $`r h2`$. 

a) What is the predicted breeding value $\hat{a_i}$ for Elsa for weaning weight? 
b) Compute the reliability of $\hat{a_i}$ from 1a) 

<!--   You can use the assigned variables from above for your computations.     --
  --   The meaning of the variables is listed below.                            --
  --      y:    phenotypic own performance record for weaning weight            --
  --      mu:   population mean for weaning weight                              --
  --      h2:   heritability for weaning weight                                 --
  --   
  --   You can use the following template of an R-code-junk for your            --
  --   computations.                                                            --ß

```{r ComputeBreedingValue, echo = TRUE}

```

  -->
```{r ResultComputeBreedingValue}
### # Compute predicted breeding value in R
hat_a_el <- h2*(y-mu)
```
  
  
### Solution

a) The predicted breeding value based on a single own performance record is

\begin{equation}
\hat{a} = h^2(y - \mu) \notag
\end{equation}



Plug-in the computed values into the equation.

\begin{equation}
\hat{a} = h^2(y - \mu) = `r h2` * (`r y`\ \text{kg} - `r mu`\ \text{kg}) = `r hat_a_el`\ \text{kg} \notag
\end{equation}

b) The reliability corresponds to the squared correlation between the true breeding value and the selection criterion. 

\begin{equation}
B = r_{a,y}^2 = h^2 = `r h2`\notag
\end{equation}



## Problem 2: Breeding Value Prediction Based on Repeated Observations
```{r SetupEx2}
geb_gew <- 52
mu2 <- 170
rep <- 0.65
```
Besides the weaning weight Elsa has also an observation for her birth weight ($`r geb_gew`$ kg) and some more repeated measures for her weight. We assume the heritability ($h^2 = `r h2`$) to be the same as in Problem 1. The population mean for the repeated observations of the weight is  $`r mu2`$ kg. The repeatability of the weight measurements is $t = `r rep`$. 

The following tables contains all observed values for the weight. 

```{r WeightTable}
nr_measure <- 10
wean_weight <- y
slope <- (wean_weight-geb_gew)/(nr_measure-1)
measure <- c(1:nr_measure)
weight <- round(slope*(measure-1) + geb_gew, digits = 0)
mean_weight <- mean(weight)
dfWeightTable <- data.frame(Measurement = measure, Weight = weight)
knitr::kable(dfWeightTable)
```

a) Predict the breeding value for Elsa based on the repeated weight records.
b) What is the reliability for the predicted breeding value from 2a)?
c) Compare the reliability from 2b) with the reliability from Problem 1.

<!--    As in Problem 1, you can use the assigned variables.           --
  --    The variables have the following meaning                       --
  --       nr_measure:   number of measurements of the weight          --
  --       h2:           heritability                                  --
  --       rep:          repeatability                                 --
  --       mean_weight:  mean of the weights                           --
  --       mu2:          population mean                               --
  -->

### Solution

a) The predicted breeding value based on repeated records is
```{r ResultHatARepMeas}
hat_a_rep_meas <- round((nr_measure * h2)/(1+(nr_measure - 1)*rep)*(mean_weight - mu2), digits = 2)
```
$$\hat{a}_i = \frac{nh^2}{1+(n-1)t}(\bar{y}_i - \mu) 
            = \frac{`r nr_measure`*`r h2`}{1+(`r nr_measure-1`*`r rep`)}(`r mean_weight` - `r mu2`)
            = `r hat_a_rep_meas`$$

b) The reliability for the predicted breeding value from 2a) is
```{r relreprecord}
rel_rep_rec <- (nr_measure * h2)/(1+(nr_measure - 1)*rep)
```


$$B = r_{a,\bar{y}}^2  = b = \frac{nh^2}{1+(n-1)t} 
                 = \frac{`r nr_measure`*`r h2`}{1+(`r nr_measure-1`*`r rep`)}
                 = `r round(rel_rep_rec, digits = 2)`$$
                 
c) The reliability of the predicted breeding values based on repeated records is larger than the reliability of the prediction based on one record. The relation between the two reliabilities is 

$$ \frac{r_{a,\bar{y}}^2}{r_{a,y}^2} = \frac{n}{1+(n-1)t} = \frac{`r nr_measure`}{1+(`r nr_measure-1`*`r rep`)}
  = `r round(nr_measure/(1+(nr_measure-1)*rep), digits = 2)`$$


## Problem 3: Predict Breeding Values based on progeny records
```{r prob3-def}
n_nr_progeny <- 5
```

A few years later Elsa was the dam of `r n_nr_progeny` offspring. Each of the offspring has a record for birthweight. Predict the breeding value of Elsa for birthweight based on the offpsring records listed in the following table.

```{r tab-prog-bw, echo=FALSE}
vec_bw_prog <- round(geb_gew + rnorm(n_nr_progeny, mean = 0, sd = 2), digits = 0)
tbl_prog_bw <- tibble::tibble(Offspring = c(1:n_nr_progeny),
                              Birthweight = vec_bw_prog)
knitr::kable(tbl_prog_bw)
```

The mean and the heritability can be taken the same as in Problems 1 and 2 resulting in 

$h^2 = $ and $\mu = $


### Solution 
The predicted breeding value based on progeny records is defined as

\begin{equation}
\hat{u_i} = b * (\bar{y_i} - \mu)
\end{equation}

where $b$ is the regression coefficient which results in 

$$b = \frac{2n}{n + k}$$

