---
title: Livestock Breeding and Genomics - Solution 4
author: Peter von Rohr
date: 2019-10-18
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'asis')
```

## Problem 1: Prediction of Breeding Value and Reliability
```{r SetUpEx1}
y <- 320
mu <- 250
h2 <- 0.45
```

Heifer Elsa has a weaning weight of $`r y`$ kg. The population mean of the same trait is at $`r mu`$ kg.The heritability ($h^2$) of the trait weaning weight is assumed to be $`r h2`$. 

a) What is the predicted breeding value $\hat{a_i}$ for Elsa for weaning weight? 
b) Compute the reliability of $\hat{a_i}$ from 1a) 

<!--   You can use the assigned variables from above for your computations.     --
  --   The meaning of the variables is listed below.                            --
  --      y:    phenotypic own performance record for weaning weight            --
  --      mu:   population mean for weaning weight                              --
  --      h2:   heritability for weaning weight                                 --
  --   
  --   You can use the following template of an R-code-junk for your            --
  --   computations.                                                            --ß

```{r ComputeBreedingValue, echo = TRUE}

```

  -->
```{r ResultComputeBreedingValue}
### # Compute predicted breeding value in R
hat_a_el <- h2*(y-mu)
```
  
  
### Solution

a) The predicted breeding value based on a single own performance record is

\begin{equation}
\hat{a} = h^2(y - \mu) \notag
\end{equation}



Plug-in the computed values into the equation.

\begin{equation}
\hat{a} = h^2(y - \mu) = `r h2` * (`r y`\ \text{kg} - `r mu`\ \text{kg}) = `r hat_a_el`\ \text{kg} \notag
\end{equation}

b) The reliability corresponds to the squared correlation between the true breeding value and the selection criterion. 

\begin{equation}
B = r_{a,y}^2 = h^2 = `r h2`\notag
\end{equation}



## Problem 2: Breeding Value Prediction Based on Repeated Observations
```{r SetupEx2}
geb_gew <- 52
mu2 <- 170
rep <- 0.65
```
Besides the weaning weight Elsa has also an observation for her birth weight ($`r geb_gew`$ kg) and some more repeated measures for her weight. We assume the heritability ($h^2 = `r h2`$) to be the same as in Problem 1. The population mean for the repeated observations of the weight is  $`r mu2`$ kg. The repeatability of the weight measurements is $t = `r rep`$. 

The following tables contains all observed values for the weight. 

```{r WeightTable}
nr_measure <- 10
wean_weight <- y
slope <- (wean_weight-geb_gew)/(nr_measure-1)
measure <- c(1:nr_measure)
weight <- round(slope*(measure-1) + geb_gew, digits = 0)
mean_weight <- mean(weight)
dfWeightTable <- data.frame(Measurement = measure, Weight = weight)
knitr::kable(dfWeightTable)
```

a) Predict the breeding value for Elsa based on the repeated weight records.
b) What is the reliability for the predicted breeding value from 2a)?
c) Compare the reliability from 2b) with the reliability from Problem 1.

<!--    As in Problem 1, you can use the assigned variables.           --
  --    The variables have the following meaning                       --
  --       nr_measure:   number of measurements of the weight          --
  --       h2:           heritability                                  --
  --       rep:          repeatability                                 --
  --       mean_weight:  mean of the weights                           --
  --       mu2:          population mean                               --
  -->

### Solution

a) The predicted breeding value based on repeated records is
```{r ResultHatARepMeas}
hat_a_rep_meas <- round((nr_measure * h2)/(1+(nr_measure - 1)*rep)*(mean_weight - mu2), digits = 2)
```
$$\hat{a}_i = \frac{nh^2}{1+(n-1)t}(\bar{y}_i - \mu) 
            = \frac{`r nr_measure`*`r h2`}{1+(`r nr_measure-1`*`r rep`)}(`r mean_weight` - `r mu2`)
            = `r hat_a_rep_meas`$$

b) The reliability for the predicted breeding value from 2a) is
```{r relreprecord}
rel_rep_rec <- (nr_measure * h2)/(1+(nr_measure - 1)*rep)
```


$$B = r_{a,\bar{y}}^2  = b = \frac{nh^2}{1+(n-1)t} 
                 = \frac{`r nr_measure`*`r h2`}{1+(`r nr_measure-1`*`r rep`)}
                 = `r round(rel_rep_rec, digits = 2)`$$
                 
c) The reliability of the predicted breeding values based on repeated records is larger than the reliability of the prediction based on one record. The relation between the two reliabilities is 

$$ \frac{r_{a,\bar{y}}^2}{r_{a,y}^2} = \frac{n}{1+(n-1)t} = \frac{`r nr_measure`}{1+(`r nr_measure-1`*`r rep`)}
  = `r round(nr_measure/(1+(nr_measure-1)*rep), digits = 2)`$$


## Problem 3: Predicted Breeding Value and Confidence Interval
During the lecture, the concept of __reliability__ ($B$) was introduced as a measure for the accuracy of a breeding value. The reliability corresponds to the squared correlation coefficient of the true breeding value ($a$) and the selection criterion ($y$). Hence 

$$B = r_{a,y}^2$$

An alternative to the reliability, we can also use the standard deviation $\sqrt{var(\hat{a})}$ of the predicted breeding value to assess the accuracy of a predicted breeding value. This standard deviation of the predicted breeding values can be used to compute confidence intervals. These intervals indicate how accurate the predicted breeding value is. A narrow confidence interval means that the predicted breeding value has a high accuracy. Each confidence interval is associated with an error probability. A very common value of an error probability is $\alpha = 0.05$. This leads to a $95\%$-confidence interval.

For a predicted breeding value $\hat{a}$ the following procedure can be used to compute the $95\%$-confidence interval. So far, we have used the regression method to predict breeding values. With that method the predicted breeding value is expressed as a function of the phenotypic observations ($y$). In our model that links the true breeding values to the phenotypic observations, we have seen that the phenotypes ($y$) are random variables which follow a certain distribution. The normal distribution is a reasonable assumption for the distribution of the phenotypic observations. As a consequence the predicted breeding values follow also a normal distribution. This distribution has its center at the predicted breeding value and the variance ($var(\hat{a})$) can be computed with the prediction equation for $a$. 

```{r DefinePhenStd}
phen_sd <- 20
```

In problem 1 where the breeding value was predicted based on an own performance records, the standard deviation $\sqrt{var(\hat{a})}$ of the predicted breeding value can be computed as follows. We assume the phenotypic standard deviation for weaning weight is `r phen_sd` kg.

```{r ResultHatAElSd}
hat_a_el_sd <- h2*phen_sd
```

\begin{equation}
\sqrt{var(\hat{a})} = \sqrt{var(b(y-\mu))} = b \sqrt{var(y)} = h^2 \sqrt{var(y)}
  = `r h2` * `r phen_sd` = `r hat_a_el_sd`\notag
\label{eq:VarHatA}
\end{equation}

Based on the properties of the normal distribution 95% of the probability mass are within $\pm 1.96$ standard deviations around the expected value. Therefore the lower ($ug$) and the upper ($og$) limits of the $95\%$-confidence interval are defined as 

```{r ConvIntHatAEl}
conv_int_hat_a_el_ug <- hat_a_el - 1.96 * hat_a_el_sd
conv_int_hat_a_el_og <- hat_a_el + 1.96 * hat_a_el_sd
```

$$ug = \hat{a} - 1.96 * \sqrt{var(\hat{a})} 
     = `r hat_a_el`\ \text{kg} - 1.96 * `r hat_a_el_sd`\ \text{kg}
     = `r conv_int_hat_a_el_ug`\ \text{kg}$$

$$og = \hat{a} + 1.96 * \sqrt{var(\hat{a})} 
     = `r hat_a_el`\ \text{kg} + 1.96 * `r hat_a_el_sd`\ \text{kg}
     = `r conv_int_hat_a_el_og`\ \text{kg}$$
     

## Your Task
Compute the $95\%$-confidence interval for the predicted breeding values in Problem 2.

<!--    You can use the assigned variables from Problem 2.                                   --
  --    First, you have to compute the standard deviation of the predicted breeding values.  --
  --    The standard deviation determines the lower and the upper bounds of the confidence   --
  --    interval.
  -->
  
  
### Solution 
For the computation of the confidence interval, we need the standard deviation of the predicted breeding value. For the example with the repeated measure, this corresponds to

```{r ResultHatARepMeasSd}
hat_a_rep_meas_sd <- nr_measure * h2/(1+(nr_measure-1)*rep) *  
    sqrt(rep + (1-rep)/nr_measure) * phen_sd
hat_a_rep_meas_sd <- round(hat_a_rep_meas_sd, digits = 2)
```

\begin{align}
\sqrt{var(\hat{a})} &= \sqrt{var(b(\bar{y}-\mu))} = b * \sqrt{var(\bar{y})} \notag\\
  &= \frac{nh^2}{1+(n-1)t} * \sqrt{\left[ t + (1-t)/n \right]\ var(y)}\notag\\
  &= \frac{`r nr_measure`*`r h2`}{1+(`r nr_measure-1`*`r rep`)} 
    * \sqrt{\left[ `r rep` + (1-`r rep`)/`r nr_measure` \right]} * `r phen_sd` \notag\\
  &= `r hat_a_rep_meas_sd`  \notag
\end{align}

The lower and the upper limits of the confidence interval are

$$ug = \hat{a} - 1.96 * \sqrt{var(\hat{a})} = `r hat_a_rep_meas` - 1.96 * `r hat_a_rep_meas_sd`
     = `r hat_a_rep_meas - 1.96 * hat_a_rep_meas_sd`$$

$$og = \hat{a} + 1.96 * \sqrt{var(\hat{a})} = `r hat_a_rep_meas` + 1.96 * `r hat_a_rep_meas_sd`
     = `r hat_a_rep_meas + 1.96 * hat_a_rep_meas_sd`$$


## Problem 4: Response To Selection
```{r problem4setup}
sel_int <- 1.4
sigma_tilde_y <- phen_sd * sqrt((1+(nr_measure-1)*rep)/nr_measure)
rel_rep_rec <- (nr_measure * h2)/(1+(nr_measure - 1)*rep)
sel_resp <- sel_int * rel_rep_rec * sigma_tilde_y
```

Assuming that we select for the predicted breeding value based on repeated measures, the response to that selection schema per generation can be computed as 

\begin{align}
R &= i * r_{a,\tilde{y}}^2 * \sigma_{\tilde{y}} \notag 
\end{align}

Compute the response to selection $R$ for the results that you found in Problem 2. The selection intensity $i$ can be assumed to be $`r sel_int`$ and the phenotypic standard deviation is taken to be the same value as in Problem 3, $\sigma_y = `r phen_sd`$.


### Solution
We insert the compute value into the equation for $R$. We have to remember that 

\begin{align}
var(\tilde{y}) &= t * \sigma_y^2 + {1\over n} (1-t) * \sigma_y^2 \notag \\
               &= {1\over n}\left( n*t + (1-t) \right) \sigma_y^2 \notag \\
               &= \frac{1 + (n-1)t}{n} \sigma_y^2 \notag
\end{align}

and $\sigma_{\tilde{y}} = \sqrt{var(\tilde{y})} = \sigma_y * \sqrt{\frac{1 + (n-1)t}{n}} = `r phen_sd` * \sqrt{\frac{1 + (`r nr_measure`-1)*`r rep`}{`r nr_measure`}} = `r round(sigma_tilde_y, digits = 2)`$

\begin{align}
R  &=  i * r_{a,\tilde{y}}^2 * \sigma_{\tilde{y}} \notag \\
   &=  `r sel_int` * `r round(rel_rep_rec, digits=2)` * `r round(sigma_tilde_y, digits = 2)` \notag \\
   &=  `r round(sel_resp, digits=2)` \notag
\end{align}

