---
title: Livestock Breeding and Genomics - Exercise 7
author: "Peter von Rohr"
date: "2019-11-08"
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    keep_tex: false
header-includes:
 \usepackage{longtable}
 \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'asis', fig.pos = 'H')
knitr::knit_hooks$set(hook_convert_odg = rmdhelp::hook_convert_odg)
```

## Problem 1: Decompositions
Given is the following pedigree. 

```{r pedsol09p01, echo=FALSE, results='asis'}
n_nr_pedsol09p01 <- 6
tbl_pedsol09p01 <- tibble::tibble(Animal = 1:n_nr_pedsol09p01,
                                      Sire   = c(NA, NA, 1, 1, 3, 5),
                                      Dam    = c(NA, NA, 2, NA, 4, 4))
knitr::kable(tbl_pedsol09p01,
             booktabs = TRUE, 
             longtable = TRUE)
```

### Your Tasks
* Do the simple decomposition of the above pedigree
* Do the recursive decomposition of the above pedigree until only $m_i$ terms appear on the right-hand side of the decomposition.
 
### Your Solution



## Problem 2: Henderson's Rules
Compute $A^{-1}$ for the following pedigree using Henderson's rules. Verify your result with the function `pedigreemm::getAInv()`. 

```{r ped_sol07p02, echo=FALSE}
n_nr_pedsol09p02 <- 6
tbl_pedsol09p02 <- tibble::tibble(Animal = 3:n_nr_pedsol09p02,
                                      Sire   = c(1, NA, 3, 3),
                                      Dam    = c(NA, 2, 4, 4))
knitr::kable(tbl_pedsol09p02,
             booktabs = TRUE, 
             longtable = TRUE)
```


### Your Solution
