---
title: "From One Locus To The Infinitesimal Model"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::knit_hooks$set(hook_convert_odg = rmdhelp::hook_convert_odg)
```


## Disclaimer
Explaining the transition from the one locus model to the infinitesimal model is difficult. Here a few attempts and experiments are shown.


## Model Backgrounds
(Chpt 7 of Mackay1996) Genetic properties of a population are expressible in terms of allele frequencies and genotype frequencies. To establish a connection between these genetic quantities on the one hand and the differences observed in a metric character on the other hand, we have to introduce the concept of **value**. Values observed in a population are **phenotypic** values of a given individual. To analyse genetic properties of populations, we have to divide the phenotypic values into components attributable to different causes. 

### Decomposition of phenotype
The first division of phenotypic values is into components attributable to the influence of genotype and environment. The two components of value associated with genotype and environment are the **genotypic value"** (G) and the **environmental deviation** (E). Hence the model can be written as 

$$P = G + E$$

In this parametrisation (definition of effects), the mean environmental deviation is $0$ and the mean phenotypic value corresponds to the mean genotypic value which is just called the **population mean** ($\mu$).


## One Locus
We assume a bi-allelic locus with allele frequencies equal to $p$ and $q = 1-p$. The genotypic values for the three genotypes are


```{r}
tbl_gv <- tibble::tibble(Genotype = c("$G_1G_1$", "$G_1G_2$", "$G_2G_2$"),
                         `Genotypic Value` = c("$a$", "$d$", "$-a$"))
knitr::kable(tbl_gv)
```


## Concept of Breeding Value
The breeding value of a given animal is defined as two times the deviation of a large number of offspring of that animal form the population mean. 

