---
title: Livestock Breeding and Genomics - Exercise 8
author: "Peter von Rohr"
date: "2019-11-15"
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    keep_tex: false
header-includes:
 \usepackage{longtable}
 \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'asis', fig.pos = 'H')
knitr::knit_hooks$set(hook_convert_odg = rmddochelper::hook_convert_odg)
```

## Problem 1: Compute Inbreeding Coefficients
Given the following pedigree.

```{r}
n_nr_ani_ped <- 6
n_nr_parent <- 3
tbl_ped_sol10p01 <- tibble::tibble(Animal = c(1:n_nr_ani_ped),
                             Sire = c(NA, NA, 1, 3, 4, 4),
                             Dam  = c(NA, NA, NA, 2, 2, 5))
### # show the pedigree
knitr::kable( tbl_ped_sol10p01,
              booktabs = TRUE,
              longtable = TRUE )
```

### Your Task
Compute the inbreeding coefficients $F_i$ for all animals using the matrix $R$ that comes from the cholesky decomposition of the numerator relationship matrix $A$



## Problem 2: Direct Construction of $A^{-1}$
Use the pedigree from problem 1 and the computed inbreeding coefficients from problem 1 to set up the inverse numerator relationship matrix $A^{-1}$ using the general form of Henderson's rules for a pedigree with inbred animals. Compare your result using function `pedigreemm::getAInv()`.


