---
title: Livestock Breeding and Genomics - Exercise 11
author: "Peter von Rohr"
date: "2019-12-06"
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    keep_tex: false
header-includes:
 \usepackage{longtable}
 \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'asis', fig.pos = 'H')
knitr::knit_hooks$set(hook_convert_odg = rmddochelper::hook_convert_odg)
```

```{r TaskCounts, echo=FALSE}
cnt <- rmddochelper::R6ClassCount$new()
cnt$set_prefix(ps_prefix = "## Problem")
```

```{r, echo=FALSE, results='asis'}
cat(cnt$out(ps_suffix = "Multivariate BLUP Animal Model"), "\n")
```

```{r sol12p01data, echo=FALSE, results='hide'}
n_nr_trait <- 2
n_nr_founder <- 3
n_nr_animal <- 8
n_nr_observation <- n_nr_animal - n_nr_founder
tbl_data_sol12p01 <- tibble::tibble(Animal = c((n_nr_founder+1):n_nr_animal),
                                        Sex = c("Male", "Female","Female","Male","Male"),
                                        Sire = c(1,3,1,4,3),
                                        Dam = c(NA,2,2,5,6),
                                        WWG = c(4.5,2.9,3.9,3.5,5.0),
                                        PWG = c(6.8,5.0,6.8,6.0,7.5))
```

The table below contains data for pre-weaning gain (WWG) and post-weaning gain (PWG) for `r n_nr_observation` beef calves.

```{r sol12p01table, echo=FALSE, results='asis'}
knitr::kable(tbl_data_sol12p01,
             booktabs = TRUE,
             longtable = TRUE)
```

The genetic variance-covariance matrix $G_0$ between the traits is 

```{r}
mat_g0 <- matrix(data = c(20,18,18,40), nrow = n_nr_trait, byrow = TRUE)
cat("$$\n")
cat("G_0 = \\left[")
cat(paste(rmddochelper::sConvertMatrixToLaTexArray(pmatAMatrix = mat_g0, pnDigits = 0), collapse = "\n"), "\n")
cat("\\right]\n")
cat("$$\n")
```

The residual variance-covariance matrix $R_0$ between the traits is

```{r}
mat_r0 <- matrix(data = c(40,11,11,30), nrow = n_nr_trait, byrow = TRUE)
cat("$$\n")
cat("R_0 = \\left[")
cat(paste(rmddochelper::sConvertMatrixToLaTexArray(pmatAMatrix = mat_r0, pnDigits = 0), collapse = "\n"), "\n")
cat("\\right]\n")
cat("$$\n")
```


### Your Task
Set up the mixed model equations for a multivariate BLUP analysis and compute the estimates for the fixed effects and the predictions for the breeding values.


### Your Solution

\vspace{3ex}



```{r, echo=FALSE, results='asis'}
cat(cnt$out(ps_suffix = "Comparison of Reliabilites"), "\n")
```

Compare the predicted breeding values and the reliabilites obtained as results of Problem 1 with results from two univariate analyses for the same traits are used in Problem 1. All parameters can be taken from Problem 1.


### Your Solution
