---
title:  Livestock Breeding and Genomics
author: Peter von Rohr
date:   16.09.2019
output: 
  html_document:
    toc_float: true
    toc: true
    theme: simplex
    highlight: default
---

Welcome to the website of   

- [751-6305-00L 	Livestock Breeding and Genomics](http://www.vorlesungsverzeichnis.ethz.ch/Vorlesungsverzeichnis/lerneinheit.view?lang=en&lerneinheitId=131686&semkez=2019W&ansicht=KATALOGDATEN&)


## Topics for Master Theses
Topics for Master Theses at Qualitas AG together with ETH are shown [here](https://charlotte-ngs.github.io/LBGFS2019/misc/20190916_master_thesis_topics/MasterThesisTopics_FS2019.html).


## Description
Methods for analysing livestock data, in particular for the estimation of breeding values: principles of selection index, introduction to BLUP, application of common models used, relationship matrix, methods for the estimation of variance components, basics of breeding programs. The material will be illustrated via exercises and assignments.



## Information
- Lecturer: Peter von Rohr
- Date: Friday 9-12
- Room: [LFW C 11](http://www.mapsearch.ethz.ch/map/map.do?gebaeudeMap=LFW&lang=en)
- Exam: 
    + Time: 09:15 - 11:15
    + Room: LFW C 11
    
### Exercise Platform
The exercise platform can be used to work on the exercises. The platform is available at: https://1-htz.quagzws.com/rstudio.

The username corresponds to the part of your e-mail address before the `@`-sign. The password is sent to you by a separate e-mail.

## Questions
Questions can be asked during the lecture and during the exercise hour or via e-mail: 

- Peter von Rohr (peter.vonrohr `at` gmail.com)

--------

```{r TablesSetup, echo=FALSE, results='hide'}
### # header names of tables
vecTableHeaders <- c("Week", "Date", "Topic")
```

## Lecture Program
```{r CourseProgram, echo=FALSE, results='asis'}
# define course start date
dCourseStart <- as.Date("2019/09/20")
# set number of weeks in semester
nNrSemesterWeeks <- 14
# define columns for weaks, dates, and subjects
Week <- 1:nNrSemesterWeeks
Date <- format(seq(dCourseStart, by = "week", length.out = nNrSemesterWeeks), "%d.%m")
Topic <- vector(mode = "character", length = nNrSemesterWeeks)
# subjects per week
Topic[which(Date == format(as.Date("2019/09/20"), "%d.%m"))] <- "Introduction to Livestock Breeding and Genomics"
Topic[which(Date == format(as.Date("2019/09/27"), "%d.%m"))] <- "Quantitative Genetics/Single Locus"
Topic[which(Date == format(as.Date("2019/10/04"), "%d.%m"))] <- "Genetic Evaluation with Different Sources of Information"
Topic[which(Date == format(as.Date("2019/10/11"), "%d.%m"))] <- "Genetic Covariance Between Relatives"
Topic[which(Date == format(as.Date("2019/10/18"), "%d.%m"))] <- "Best Linear Unbiased Prediction - Univariate Analysis"
Topic[which(Date == format(as.Date("2019/10/25"), "%d.%m"))] <- "Best Linear Unbiased Prediction - Multivariate Analysis"
Topic[which(Date == format(as.Date("2019/11/01"), "%d.%m"))] <- "Models with Random Environmental Effects"
Topic[which(Date == format(as.Date("2019/11/08"), "%d.%m"))] <- "Analysis of Longitudinal Data"
Topic[which(Date == format(as.Date("2019/11/15"), "%d.%m"))] <- "Variance Components Estimation"
Topic[which(Date == format(as.Date("2019/11/22"), "%d.%m"))] <- "Linkage Disequilibrium"
Topic[which(Date == format(as.Date("2019/11/29"), "%d.%m"))] <- "Genomic Selection"
Topic[which(Date == format(as.Date("2019/12/06"), "%d.%m"))] <- "Genom-Wide Association Studies"
Topic[which(Date == format(as.Date("2019/12/13"), "%d.%m"))] <- "Questions, Test Exam"
Topic[which(Date == format(as.Date("2019/12/20"), "%d.%m"))] <- "Exam"

dfCnTable <- data.frame(Week, Date, Topic, stringsAsFactors = FALSE)
colnames(dfCnTable) <- vecTableHeaders
knitr::kable(dfCnTable)
```

--------

## Material
Course material (notes and slides) and Exercises with Solutions can be downloaded from this section. 

### Course Notes
```{r MaterialTable, echo=FALSE, results='asis'}
# define course start date
dCourseStart <- as.Date("2019/09/20")
# set number of weeks in semester
nNrSemesterWeeks <- 14
# define columns for weaks, dates, and subjects
Woche <- 1:nNrSemesterWeeks
Date <- format(seq(dCourseStart, by = "week", length.out = nNrSemesterWeeks), "%d.%m")
vecCnTopic <- vector(mode = "character", length = nNrSemesterWeeks)
# exam
vecCnTopic[length(vecCnTopic)] <- "Exam"
# base url
s_base_url_cn <- 'https://charlotte-ngs.github.io/LBGFS2019/cn_lbg'
# subjects per week
### # w1
vecCnTopic[which(Date == format(as.Date("2019/09/20"), "%d.%m"))] <- paste0("[Introduction to Livestock Breeding and Genomics](", s_base_url_cn, "/03_intro_lbg.pdf)", collapse = '')
### # w2
vecCnTopic[which(Date == format(as.Date("2019/09/27"), "%d.%m"))] <- paste0("[Quantitative Genetics](", s_base_url_cn, "/04_basics_qg.pdf)", collapse = '')
### # w3
vecCnTopic[which(Date == format(as.Date("2019/10/04"), "%d.%m"))] <- paste0("[Genetic Evaluation](", s_base_url_cn, "/05_gen_eval.pdf)", collapse = '')
### # w4 cn/03-prediction-bv.pdf
vecCnTopic[which(Date == format(as.Date("2019/10/11"), "%d.%m"))] <- paste0("[BLUP](", s_base_url_cn, "/06_blup.pdf)", collapse = '')
### # w5 04-selection-index.pdf
vecCnTopic[which(Date == format(as.Date("2019/10/18"), "%d.%m"))] <- paste0("[Genetic Covariance](", s_base_url_cn, "/07_gen_cov.pdf)", collapse = '')
### # w6 cn/05-blup.pdf
vecCnTopic[which(Date == format(as.Date("2019/10/25"), "%d.%m"))] <- paste0("[Additional Aspects of BLUP](", s_base_url_cn, "/08_add_asp_blup.pdf)", collapse = '')
### # w7 cn/06-gencov.pdf
# vecCnTopic[which(Date == format(as.Date("2019/11/01"), "%d.%m"))] <- "[]()"
### # w8 cn/063-matainv.pdf
# vecCnTopic[which(Date == format(as.Date("2019/11/08"), "%d.%m"))] <- "[]()"
### # w9 cn/07-add-asp-blup.pdf
# vecCnTopic[which(Date == format(as.Date("2019/11/15"), "%d.%m"))] <- "[]()"
### # w10 08-multiple-traits.pdf
# vecCnTopic[which(Date == format(as.Date("2019/11/22"), "%d.%m"))] <- "[]()"
### # w11 09-genomic-selection.pdf
# vecCnTopic[which(Date == format(as.Date("2019/11/29"), "%d.%m"))] <- "[]()"
### # w12 10-var-comp.pdf
# vecCnTopic[which(Date == format(as.Date("2019/12/06"), "%d.%m"))] <- "[]()"

dfCnTable <- data.frame(Woche, Date, vecCnTopic, stringsAsFactors = FALSE)
colnames(dfCnTable) <- vecTableHeaders
knitr::kable(dfCnTable)
```


### Slides
```{r SlidesTable, echo=FALSE, results='asis'}
### # define vector for subjects
vecSlTopic = vector(mode = "character", length = nNrSemesterWeeks)
# exam
vecSlTopic[length(Topic)] <- "Exam"
# base url
s_base_url_sl <- 'https://charlotte-ngs.github.io/LBGFS2019/sl'
### # assign slides
### # w1 
vecSlTopic[which(Date == format(as.Date("2019/09/20"), "%d.%m"))] <- paste0("[Introduction to Livestock Breeding and Genomics](", s_base_url_sl, "/w01/lbgfs2019_slides_w01_intro-lbg.pdf)", collapse = '')
### # w2
vecSlTopic[which(Date == format(as.Date("2019/09/27"), "%d.%m"))] <- paste0("[Quantitative Genetics](", s_base_url_sl,"/w02/lbg_fs_2019_slides_basics-qg.pdf), [Part 2](", s_base_url_sl, "/w03/lbgfs2019_slides_basics-qg-teil2.pdf)", collapse = '')
### # w3
vecSlTopic[which(Date == format(as.Date("2019/10/04"), "%d.%m"))] <- paste0("[Genetic Evaluation](", s_base_url_sl,"/w04/lbgfs2019_slides_gen-eval.pdf)", collapse = '')
### # w4 w04/lbg_fs_2019_slides_gen-eval.pdf
vecSlTopic[which(Date == format(as.Date("2019/10/11"), "%d.%m"))] <- paste0("[BLUP](", s_base_url_sl,"/w05/lbgfs2019_slides_blup.pdf)", collapse = '')
### # w5 w05/lbg_fs_2019_slides_selection-index.pdf
vecSlTopic[which(Date == format(as.Date("2019/10/18"), "%d.%m"))] <- paste0("[Genetic Covariance](", s_base_url_sl,"/w06/lbgfs2019_slides_gencov.pdf)", collapse = '')
### # w6 w06/lbg_fs_2019_slides_blup.pdf
vecSlTopic[which(Date == format(as.Date("2019/10/25"), "%d.%m"))] <- paste0("[Numerator Relationship Matrix](", s_base_url_sl,"/w07/lbgfs2019_slides_numrelmat.pdf)", collapse = '')
### # w7 w07/lbg_fs_2019_slides_pbvsi.pdf w07/lbg_fs_2019_slides_gencov.pdf
# vecSlTopic[which(Date == format(as.Date("2019/11/01"), "%d.%m"))] <- "[]()"
### # w8 w08/lbg_fs_2019_slides_numrelmat.pdf
# vecSlTopic[which(Date == format(as.Date("2019/11/08"), "%d.%m"))] <- "[]()"
### # w9 w09/lbg_fs_2019_slides_invnumrelmat.pdf
# vecSlTopic[which(Date == format(as.Date("2019/11/15"), "%d.%m"))] <- "[]()"
### # w10 /w10/lbg_fs_2019_slides_inbreeding.pdf
# vecSlTopic[which(Date == format(as.Date("2019/11/22"), "%d.%m"))] <- "[]()"
### # w11 w11/lbg_fs_2019_slides_additional_aspects.pdf
# vecSlTopic[which(Date == format(as.Date("2019/11/29"), "%d.%m"))] <- "[]()"
### # w12
# vecSlTopic[which(Date == format(as.Date("2019/12/06"), "%d.%m"))] <- "[]()"

### # construct the table
dfSlTable <- data.frame(Woche, Date, vecSlTopic, stringsAsFactors = FALSE)
colnames(dfSlTable) <- vecTableHeaders
knitr::kable(dfSlTable)
```


### Exercises
```{r ExerciseTable, echo=FALSE, results='asis'}
### # define vector for subjects
vecExTopic = vector(mode = "character", length = nNrSemesterWeeks)
# set subject for special dates
# exam
vecExTopic[length(Topic)] <- "Exam"
# base url
s_base_url_ex <- 'https://charlotte-ngs.github.io/LBGFS2019/ex'
# w2
vecExTopic[which(Date == format(as.Date("2019/09/27"), "%d.%m"))] <- paste0("[Introduction to Linear Algebra and Livestock Breeding](", s_base_url_ex, "/w02/lbgfs2019_w02_ex01.pdf)", collapse = '')
# w3
vecExTopic[which(Date == format(as.Date("2019/10/04"), "%d.%m"))] <- paste0("[Matrices and Quantitative Genetics](", s_base_url_ex, "/w03/lbgfs2019_w03_ex02.pdf)", collapse = '')
# w4 w04/lbg_fs_2019_w04_ex3.pdf
vecExTopic[which(Date == format(as.Date("2019/10/11"), "%d.%m"))] <- paste0("[Single Locus Breeding Values](", s_base_url_ex, "/w04/lbgfs2019_w04_ex03.pdf)", collapse = '')
# w5 ex/w05/lbg_fs_2019_w05_ex04.pdf
vecExTopic[which(Date == format(as.Date("2019/10/18"), "%d.%m"))] <- paste0("[Own Performance and Repeated Records](", s_base_url_ex, "/w05/lbgfs2019_w05_ex04.pdf)", collapse = '')
# w6 ex/w06/lbg_fs_2019_w06_ex05.pdf
vecExTopic[which(Date == format(as.Date("2019/10/25"), "%d.%m"))] <- paste0("[Progeny Records and Sire Model](", s_base_url_ex, "/w06/lbgfs2019_w06_ex05.pdf)", collapse = '')
# w7  w07/lbg_fs_2019_w07_ex06.pdf
# vecExTopic[which(Date == format(as.Date("2019/11/01"), "%d.%m"))] <- "[]()"
# w8 ex/w08/lbg_fs_2019_w08_ex07.pdf
# vecExTopic[which(Date == format(as.Date("2019/11/08"), "%d.%m"))] <- "[]()"
# w9 ex/w09/lbg_fs_2019_w09_ex08.pdf
# vecExTopic[which(Date == format(as.Date("2019/11/15"), "%d.%m"))] <- "[]()"
# w10 w10/lbg_fs_2019_w10_ex09.pdf
# vecExTopic[which(Date == format(as.Date("2019/11/22"), "%d.%m"))] <- "[]()"
# w11 ex/w11/lbg_fs_2019_w11_ex10.pdf 
# vecExTopic[which(Date == format(as.Date("2019/11/29"), "%d.%m"))] <- "[]()"
# w12 ex/w12/lbg_fs_2019_w12_ex11.pdf
# vecExTopic[which(Date == format(as.Date("2019/12/06"), "%d.%m"))] <- "[]()"
# w13 w13/lbg_hs2017_exam.pdf
# vecExTopic[which(Date == format(as.Date("2019/12/13"), "%d.%m"))] <- "[]()"

### # generate table output
dfExTable <- data.frame(Woche, Date, vecExTopic, stringsAsFactors = FALSE)
colnames(dfExTable) <- vecTableHeaders
knitr::kable(dfExTable)
```


### Solution
```{r SolutionTable, echo=FALSE, results='asis'}
### # define vector for subjects
vecSolTopic = vector(mode = "character", length = nNrSemesterWeeks)
# base url
s_base_url_sol <- 'https://charlotte-ngs.github.io/LBGFS2019/sol'
# w3
vecSolTopic[which(Date == format(as.Date("2019/10/04"), "%d.%m"))] <- paste0("[Introduction to Linear Algebra and Livestock Breeding](", s_base_url_sol, "/w02/lbgfs2019_w02_sol01.pdf)", collapse = '')
# w4 Matrices and Population Mean w03/lbg_fs_2019_w03_sol02.pdf
vecSolTopic[which(Date == format(as.Date("2019/10/11"), "%d.%m"))] <- paste0("[Matrices and Quantitative Genetics](", s_base_url_sol, "/w03/lbgfs2019_w03_sol02.pdf)", collapse = '')
# w5 sol/w04/lbg_fs_2019_w04_sol3.pdf
vecSolTopic[which(Date == format(as.Date("2019/10/18"), "%d.%m"))] <- paste0("[Single Locus Breeding Values](", s_base_url_sol, "/w04/lbgfs2019_w04_sol03.pdf)", collapse = '')
# w6 sol/w05/lbg_fs_2019_w05_sol04.pdf
vecSolTopic[which(Date == format(as.Date("2019/10/25"), "%d.%m"))] <- paste0("[Own Performance and Repeated Records](", s_base_url_sol, "/w05/lbgfs2019_w05_sol04.pdf)", collapse = '')
# w7 sol/w06/lbg_fs_2019_w06_sol05.pdf
# vecSolTopic[which(Date == format(as.Date("2019/11/01"), "%d.%m"))] <- paste0("[Progeny Records and Sire Model](", s_base_url_sol, "/w06/lbgfs2019_w06_sol05.pdf)", collapse = '')
# w8 w07/lbg_fs_2019_w07_sol06.pdf
# vecSolTopic[which(Date == format(as.Date("2019/11/08"), "%d.%m"))] <- "[]()"
# w9 sol/w08/lbg_fs_2019_w08_sol07.pdf
# vecSolTopic[which(Date == format(as.Date("2019/11/15"), "%d.%m"))] <- "[]()"
# w10 sol/w09/lbg_fs_2019_w09_sol08.pdf
# vecSolTopic[which(Date == format(as.Date("2019/11/22"), "%d.%m"))] <- "[]()"
# w11 w10/lbg_fs_2019_w10_sol09.pdf
# vecSolTopic[which(Date == format(as.Date("2019/11/29"), "%d.%m"))] <- "[]()"
# w12 sol/w11/lbg_fs_2019_w11_sol10.pdf
# vecSolTopic[which(Date == format(as.Date("2019/12/06"), "%d.%m"))] <- "[]()"
# w13 sol/w12/lbg_fs_2019_w12_sol11.pdf
# vecSolTopic[which(Date == format(as.Date("2019/12/13"), "%d.%m"))] <- "[]()"
# w14 
# vecSolTopic[which(Date == format(as.Date("2019/12/20"), "%d.%m"))] <- "[]()"

### # construct the table
dfSolTable <- data.frame(Woche, Date, vecSolTopic, stringsAsFactors = FALSE)
### # generate table output
colnames(dfSolTable) <- vecTableHeaders
knitr::kable(dfSolTable)
```


### More Material
```{r FurtherMatTable, echo=FALSE, results='asis'}
### # define vector for subjects
vecAddMatTopic = vector(mode = "character", length = nNrSemesterWeeks)
### # w1
vecAddMatTopic[which(Date == format(as.Date("2019/09/20"), "%d.%m"))] <- "[Introduction to Linear Algebra](https://charlotte-ngs.github.io/LBGFS2019/cn_lbg/01_intro_linalg.pdf), [Introduction to R/RStudio](https://charlotte-ngs.github.io/LBGFS2019/cn_lbg/02_intro_rstats.pdf), [Questionnaire](https://charlotte-ngs.github.io/LBGFS2019/misc/20190920_students_questionnaire/20190920_students_questionnaire.pdf)"

### # construct the table
dfAddMatTable <- data.frame(Woche, Date, vecAddMatTopic, stringsAsFactors = FALSE)

### # generate table output
colnames(dfAddMatTable) <- vecTableHeaders
knitr::kable(dfAddMatTable)
```

```{r LatestChange, echo=FALSE, results='asis'}
cat(rmdhelp::get_latest_change(), "\n")
```


