---
title:  Livestock Breeding and Genomics
author: Peter von Rohr
date:   16.09.2019
output: 
  html_document:
    toc_float: true
    toc: true
    theme: paper
    highlight: default
---

Welcome to the website of   

- [751-6305-00L 	Livestock Breeding and Genomics](http://www.vorlesungsverzeichnis.ethz.ch/Vorlesungsverzeichnis/lerneinheit.view?lang=en&lerneinheitId=131686&semkez=2019W&ansicht=KATALOGDATEN&)


## Topics for Master Theses
Topics for Master Theses at Qualitas AG together with ETH are shown [here](https://charlotte-ngs.github.io/LBGFS2018/misc/20181026_master_thesis_topics/MasterThesisTopics_FS2018.html).


## Description
Methods for analysing livestock data, in particular for the estimation of breeding values: principles of selection index, introduction to BLUP, application of common models used, relationship matrix, methods for the estimation of variance components, basics of breeding programs. The material will be illustrated via exercises and assignments.



## Information
- Lecturer: Peter von Rohr
- Date: Friday 9-12
- Room: [LFW C 11](http://www.mapsearch.ethz.ch/map/map.do?gebaeudeMap=LFW&lang=en)
- Exam: 
    + Time: 09:15 - 11:15
    + Room: LFW C 11
    
### Exercise Platform
The exercise platform can be used to work on the exercises. The platform can be reached at: http://r4tea.rteastem.org:8787/

The username corresponds to the part of your e-mail address before the `@`-sign. The password is sent to you by a separate e-mail.

## Questions
Questions can be asked during the lecture and during the exercise hour or via e-mail: 

- Peter von Rohr (peter.vonrohr `at` gmail.com)

--------

```{r TablesSetup, echo=FALSE, results='hide'}
### # header names of tables
vecTableHeaders <- c("Week", "Date", "Topic")
```

## Lecture Program
```{r CourseProgram, echo=FALSE, results='asis'}
# define course start date
dCourseStart <- as.Date("2018/09/21")
# set number of weeks in semester
nNrSemesterWeeks <- 14
# define columns for weaks, dates, and subjects
Week <- 1:nNrSemesterWeeks
Date <- format(seq(dCourseStart, by = "week", length.out = nNrSemesterWeeks), "%d.%m")
Topic <- vector(mode = "character", length = nNrSemesterWeeks)
# subjects per week
Topic[which(Date == format(as.Date("2018/09/21"), "%d.%m"))] <- "Introduction to Livestock Breeding and Genomics"
Topic[which(Date == format(as.Date("2018/09/28"), "%d.%m"))] <- "Quantitative Genetics/Single Locus"
Topic[which(Date == format(as.Date("2018/10/05"), "%d.%m"))] <- "Genetic Evaluation with Different Sources of Information"
Topic[which(Date == format(as.Date("2018/10/12"), "%d.%m"))] <- "Genetic Covariance Between Relatives"
Topic[which(Date == format(as.Date("2018/10/19"), "%d.%m"))] <- "Best Linear Unbiased Prediction - Univariate Analysis"
Topic[which(Date == format(as.Date("2018/10/26"), "%d.%m"))] <- "Best Linear Unbiased Prediction - Multivariate Analysis"
Topic[which(Date == format(as.Date("2018/11/02"), "%d.%m"))] <- "Models with Random Environmental Effects"
Topic[which(Date == format(as.Date("2018/11/09"), "%d.%m"))] <- "Analysis of Longitudinal Data"
Topic[which(Date == format(as.Date("2018/11/16"), "%d.%m"))] <- "Variance Components Estimation"
Topic[which(Date == format(as.Date("2018/11/23"), "%d.%m"))] <- "Linkage Disequilibrium"
Topic[which(Date == format(as.Date("2018/11/30"), "%d.%m"))] <- "Genomic Selection"
Topic[which(Date == format(as.Date("2018/12/07"), "%d.%m"))] <- "Genom-Wide Association Studies"
Topic[which(Date == format(as.Date("2018/12/14"), "%d.%m"))] <- "Questions, Test Exam"
Topic[which(Date == format(as.Date("2018/12/21"), "%d.%m"))] <- "Exam"

dfCnTable <- data.frame(Week, Date, Topic, stringsAsFactors = FALSE)
colnames(dfCnTable) <- vecTableHeaders
knitr::kable(dfCnTable)
```

--------

## Material
Course material (notes and slides) and Exercises with Solutions can be downloaded from this section. 

### Course Notes
```{r MaterialTable, echo=FALSE, results='asis'}
# define course start date
dCourseStart <- as.Date("2018/09/21")
# set number of weeks in semester
nNrSemesterWeeks <- 14
# define columns for weaks, dates, and subjects
Woche <- 1:nNrSemesterWeeks
Date <- format(seq(dCourseStart, by = "week", length.out = nNrSemesterWeeks), "%d.%m")
vecCnTopic <- vector(mode = "character", length = nNrSemesterWeeks)
# exam
vecCnTopic[length(vecCnTopic)] <- "Exam"
# subjects per week
### # w1
vecCnTopic[which(Date == format(as.Date("2018/09/21"), "%d.%m"))] <- "[Introduction to Livestock Breeding and Genomics](https://charlotte-ngs.github.io/LBGFS2018/cn/01-intro-lbg.pdf)"
### # w2
vecCnTopic[which(Date == format(as.Date("2018/09/28"), "%d.%m"))] <- "[Basics of Quantitative Genetics](https://charlotte-ngs.github.io/LBGFS2018/cn/02-basics-qg.pdf)"
### # w3
vecCnTopic[which(Date == format(as.Date("2018/10/05"), "%d.%m"))] <- "[Basics of Quantitative Genetics (Teil 2)](https://charlotte-ngs.github.io/LBGFS2018/cn/02-basic-qg-teil2.pdf)"

### # w4 cn/03-prediction-bv.pdf
vecCnTopic[which(Date == format(as.Date("2018/10/12"), "%d.%m"))] <- "[Genetic Evaluation](https://charlotte-ngs.github.io/LBGFS2018/cn/03-prediction-bv.pdf)"

### # w5 04-selection-index.pdf
vecCnTopic[which(Date == format(as.Date("2018/10/19"), "%d.%m"))] <- "[Selection Index](https://charlotte-ngs.github.io/LBGFS2018/cn/04-selection-index.pdf)"

### # w6 cn/05-blup.pdf
vecCnTopic[which(Date == format(as.Date("2018/10/26"), "%d.%m"))] <- "[BLUP](https://charlotte-ngs.github.io/LBGFS2018/cn/05-blup.pdf)"

### # w7 cn/06-gencov.pdf
vecCnTopic[which(Date == format(as.Date("2018/11/02"), "%d.%m"))] <- "[Genetic Covariance Between Animals](https://charlotte-ngs.github.io/LBGFS2018/cn/06-gencov.pdf)"

### # w8 cn/063-matainv.pdf
vecCnTopic[which(Date == format(as.Date("2018/11/09"), "%d.%m"))] <- "[Inverse Numerator Relationship Matrix](https://charlotte-ngs.github.io/LBGFS2018/cn/063-matainv.pdf)"

### # w9 cn/07-add-asp-blup.pdf
vecCnTopic[which(Date == format(as.Date("2018/11/16"), "%d.%m"))] <- "[Additional Aspects of BLUP](https://charlotte-ngs.github.io/LBGFS2018/cn/07-add-asp-blup.pdf)"

### # w10 08-multiple-traits.pdf
vecCnTopic[which(Date == format(as.Date("2018/11/23"), "%d.%m"))] <- "[Multiple Traits](https://charlotte-ngs.github.io/LBGFS2018/cn/08-multiple-traits.pdf)"

### # w11 09-genomic-selection.pdf
vecCnTopic[which(Date == format(as.Date("2018/11/30"), "%d.%m"))] <- "[Genomic Selection](https://charlotte-ngs.github.io/LBGFS2018/cn/09-genomic-selection.pdf)"

### # w12 10-var-comp.pdf
vecCnTopic[which(Date == format(as.Date("2018/12/07"), "%d.%m"))] <- "[Variance Components](https://charlotte-ngs.github.io/LBGFS2018/cn/10-var-comp.pdf)"



dfCnTable <- data.frame(Woche, Date, vecCnTopic, stringsAsFactors = FALSE)
colnames(dfCnTable) <- vecTableHeaders
knitr::kable(dfCnTable)
```


### Slides
```{r SlidesTable, echo=FALSE, results='asis'}
### # define vector for subjects
vecSlTopic = vector(mode = "character", length = nNrSemesterWeeks)
# exam
vecSlTopic[length(Topic)] <- "Exam"
### # assign slides
### # w1 
vecSlTopic[which(Date == format(as.Date("2018/09/21"), "%d.%m"))] <- "[Introduction to Livestock Breeding and Genomics](https://charlotte-ngs.github.io/LBGFS2018/sl/w01/lbg_fs_2018_slides_intro-lbg.pdf)"

### # w2
vecSlTopic[which(Date == format(as.Date("2018/09/28"), "%d.%m"))] <- "[Basics of Quantitative Genetics](https://charlotte-ngs.github.io/LBGFS2018/sl/w02/lbg_fs_2018_slides_basics-qg.pdf)"

### # w3
vecSlTopic[which(Date == format(as.Date("2018/10/05"), "%d.%m"))] <- "[Basics of Quantitative Genetics (Part 2)](https://charlotte-ngs.github.io/LBGFS2018/sl/w03/lbg_fs_2018_slides_basics-qg-teil2.pdf)"

### # w4 w04/lbg_fs_2018_slides_gen-eval.pdf
vecSlTopic[which(Date == format(as.Date("2018/10/12"), "%d.%m"))] <- "[Genetic Evaluation](https://charlotte-ngs.github.io/LBGFS2018/sl/w04/lbg_fs_2018_slides_gen-eval.pdf)"


### # w5 w05/lbg_fs_2018_slides_selection-index.pdf
vecSlTopic[which(Date == format(as.Date("2018/10/19"), "%d.%m"))] <- "[Selection Index](https://charlotte-ngs.github.io/LBGFS2018/sl/w05/lbg_fs_2018_slides_selection-index.pdf)"

### # w6 w06/lbg_fs_2018_slides_blup.pdf
vecSlTopic[which(Date == format(as.Date("2018/10/26"), "%d.%m"))] <- "[BLUP](https://charlotte-ngs.github.io/LBGFS2018/sl/w06/lbg_fs_2018_slides_blup.pdf)"


### # w7 w07/lbg_fs_2018_slides_pbvsi.pdf w07/lbg_fs_2018_slides_gencov.pdf
vecSlTopic[which(Date == format(as.Date("2018/11/02"), "%d.%m"))] <- "[Breeding Value Prediction Using Selection Index](https://charlotte-ngs.github.io/LBGFS2018/sl/w07/lbg_fs_2018_slides_pbvsi.pdf), [Genetic Covariance Between Animals](https://charlotte-ngs.github.io/LBGFS2018/sl/w07/lbg_fs_2018_slides_gencov.pdf)"

### # w8 w08/lbg_fs_2018_slides_numrelmat.pdf
vecSlTopic[which(Date == format(as.Date("2018/11/09"), "%d.%m"))] <- "[Numerator Relationship Matrix](https://charlotte-ngs.github.io/LBGFS2018/sl/w08/lbg_fs_2018_slides_numrelmat.pdf)"

### # w9 w09/lbg_fs_2018_slides_invnumrelmat.pdf
vecSlTopic[which(Date == format(as.Date("2018/11/16"), "%d.%m"))] <- "[Inverse Numerator Relationship Matrix](https://charlotte-ngs.github.io/LBGFS2018/sl/w09/lbg_fs_2018_slides_invnumrelmat.pdf)"

### # w10 /w10/lbg_fs_2018_slides_inbreeding.pdf
vecSlTopic[which(Date == format(as.Date("2018/11/23"), "%d.%m"))] <- "[Inbreeding](https://charlotte-ngs.github.io/LBGFS2018/sl/w10/lbg_fs_2018_slides_inbreeding.pdf)"

### # w11 w11/lbg_fs_2018_slides_additional_aspects.pdf
vecSlTopic[which(Date == format(as.Date("2018/11/30"), "%d.%m"))] <- "[Additional Aspects of BLUP](https://charlotte-ngs.github.io/LBGFS2018/sl/w11/lbg_fs_2018_slides_additional_aspects.pdf)"

### # w12
vecSlTopic[which(Date == format(as.Date("2018/12/07"), "%d.%m"))] <- "[Multiple Trait](https://charlotte-ngs.github.io/LBGFS2018/sl/w12/lbg_fs_2018_slides_multiple_traits.pdf), [Variance Components](https://charlotte-ngs.github.io/LBGFS2018/sl/w12/lbg_fs_2018_slides_var_comp.pdf)"


### # construct the table
dfSlTable <- data.frame(Woche, Date, vecSlTopic, stringsAsFactors = FALSE)
colnames(dfSlTable) <- vecTableHeaders
knitr::kable(dfSlTable)
```


### Exercises
```{r ExerciseTable, echo=FALSE, results='asis'}
### # define vector for subjects
vecExTopic = vector(mode = "character", length = nNrSemesterWeeks)
# set subject for special dates
# exam
vecExTopic[length(Topic)] <- "Exam"

# w2
vecExTopic[which(Date == format(as.Date("2018/09/28"), "%d.%m"))] <- "[Introduction to Livestock Breeding and Genomics](https://charlotte-ngs.github.io/LBGFS2018/ex/w02/lbg_fs_2018_ex01_w02.pdf)"

# w3
vecExTopic[which(Date == format(as.Date("2018/10/05"), "%d.%m"))] <- "[Matrices and Population Mean](https://charlotte-ngs.github.io/LBGFS2018/ex/W03/lbg_fs_2018_w03_ex02.pdf)"

# w4 w04/lbg_fs_2018_w04_ex3.pdf
vecExTopic[which(Date == format(as.Date("2018/10/12"), "%d.%m"))] <- "[Breeding Values, Allele Substitution and Data Input Mean](https://charlotte-ngs.github.io/LBGFS2018/ex/w04/lbg_fs_2018_w04_ex3.pdf)"

# w5 ex/w05/lbg_fs_2018_w05_ex04.pdf
vecExTopic[which(Date == format(as.Date("2018/10/19"), "%d.%m"))] <- "[Breeding Value Prediction, Confidence Interval and Selection Response](https://charlotte-ngs.github.io/LBGFS2018/ex/w05/lbg_fs_2018_w05_ex04.pdf)"

# w6 ex/w06/lbg_fs_2018_w06_ex05.pdf
vecExTopic[which(Date == format(as.Date("2018/10/26"), "%d.%m"))] <- "[Breeding Value Prediction, Own Performance and Progeny Records](https://charlotte-ngs.github.io/LBGFS2018/ex/w06/lbg_fs_2018_w06_ex05.pdf)"

# w7  w07/lbg_fs_2018_w07_ex06.pdf
vecExTopic[which(Date == format(as.Date("2018/11/02"), "%d.%m"))] <- "[Selection Index](https://charlotte-ngs.github.io/LBGFS2018/ex/w07/lbg_fs_2018_w07_ex06.pdf)"

# w8 ex/w08/lbg_fs_2018_w08_ex07.pdf
vecExTopic[which(Date == format(as.Date("2018/11/09"), "%d.%m"))] <- "[Selection Index and Sire Model](https://charlotte-ngs.github.io/LBGFS2018/ex/w08/lbg_fs_2018_w08_ex07.pdf)"

# w9 ex/w09/lbg_fs_2018_w09_ex08.pdf
vecExTopic[which(Date == format(as.Date("2018/11/16"), "%d.%m"))] <- "[Numerator Relationship Matrix and Animal Model](https://charlotte-ngs.github.io/LBGFS2018/ex/w09/lbg_fs_2018_w09_ex08.pdf)"

# w10 w10/lbg_fs_2018_w10_ex09.pdf
vecExTopic[which(Date == format(as.Date("2018/11/23"), "%d.%m"))] <- "[Decomposition and Covariance](https://charlotte-ngs.github.io/LBGFS2018/ex/w10/lbg_fs_2018_w10_ex09.pdf)"

# w11 ex/w11/lbg_fs_2018_w11_ex10.pdf 
vecExTopic[which(Date == format(as.Date("2018/11/30"), "%d.%m"))] <- "[Inbreeding and Henderson's Rule](https://charlotte-ngs.github.io/LBGFS2018/ex/w11/lbg_fs_2018_w11_ex10.pdf)"

# w12 ex/w12/lbg_fs_2018_w12_ex11.pdf
vecExTopic[which(Date == format(as.Date("2018/12/07"), "%d.%m"))] <- "[Accuracy and Decomposition of PBV](https://charlotte-ngs.github.io/LBGFS2018/ex/w12/lbg_fs_2018_w12_ex11.pdf)"
 

# w13 w13/lbg_hs2017_exam.pdf
vecExTopic[which(Date == format(as.Date("2018/12/14"), "%d.%m"))] <- "[Final Exam 2017](https://charlotte-ngs.github.io/LBGFS2018/ex/w13/lbg_hs2017_exam.pdf), [Multivariate BLUP and Variance Components](https://charlotte-ngs.github.io/LBGFS2018/ex/w13/lbg_fs_2018_w13_ex12.pdf)"


### # generate table output
dfExTable <- data.frame(Woche, Date, vecExTopic, stringsAsFactors = FALSE)
colnames(dfExTable) <- vecTableHeaders
knitr::kable(dfExTable)
```


### Solution
```{r SolutionTable, echo=FALSE, results='asis'}
### # define vector for subjects
vecSolTopic = vector(mode = "character", length = nNrSemesterWeeks)
# w2

# w3
vecSolTopic[which(Date == format(as.Date("2018/10/05"), "%d.%m"))] <- "[Introduction to Livestock Breeding and Genomics](https://charlotte-ngs.github.io/LBGFS2018/sol/w02/lbg_fs_2018_sol01_w02.pdf)"

# w4 Matrices and Population Mean w03/lbg_fs_2018_w03_sol02.pdf
vecSolTopic[which(Date == format(as.Date("2018/10/12"), "%d.%m"))] <- "[Matrices and Population Mean](https://charlotte-ngs.github.io/LBGFS2018/sol/w03/lbg_fs_2018_w03_sol02.pdf)"

# w5 sol/w04/lbg_fs_2018_w04_sol3.pdf
vecSolTopic[which(Date == format(as.Date("2018/10/19"), "%d.%m"))] <- "[Breeding Values, Allele Substitution and Data Input Mean](https://charlotte-ngs.github.io/LBGFS2018/sol/w04/lbg_fs_2018_w04_sol3.pdf)"

# w6 sol/w05/lbg_fs_2018_w05_sol04.pdf
vecSolTopic[which(Date == format(as.Date("2018/10/26"), "%d.%m"))] <- "[Breeding Value Prediction, Confidence Interval and Selection Response](https://charlotte-ngs.github.io/LBGFS2018/sol/w05/lbg_fs_2018_w05_sol04.pdf)"

# w7 sol/w06/lbg_fs_2018_w06_sol05.pdf
vecSolTopic[which(Date == format(as.Date("2018/11/02"), "%d.%m"))] <- "[Breeding Value Prediction, Own Performance and Progeny Records](https://charlotte-ngs.github.io/LBGFS2018/sol/w06/lbg_fs_2018_w06_sol05.pdf)"

# w8 w07/lbg_fs_2018_w07_sol06.pdf
vecSolTopic[which(Date == format(as.Date("2018/11/09"), "%d.%m"))] <- "[Selection Index](https://charlotte-ngs.github.io/LBGFS2018/sol/w07/lbg_fs_2018_w07_sol06.pdf)"

# w9 sol/w08/lbg_fs_2018_w08_sol07.pdf
vecSolTopic[which(Date == format(as.Date("2018/11/16"), "%d.%m"))] <- "[Selection Index and Sire Model](https://charlotte-ngs.github.io/LBGFS2018/sol/w08/lbg_fs_2018_w08_sol07.pdf)"

# w10 sol/w09/lbg_fs_2018_w09_sol08.pdf
vecSolTopic[which(Date == format(as.Date("2018/11/23"), "%d.%m"))] <- "[Numerator Relationship Matrix and Animal Model](https://charlotte-ngs.github.io/LBGFS2018/sol/w09/lbg_fs_2018_w09_sol08.pdf)"

# w11 w10/lbg_fs_2018_w10_sol09.pdf
vecSolTopic[which(Date == format(as.Date("2018/11/30"), "%d.%m"))] <- "[Decomposition and Covariance](https://charlotte-ngs.github.io/LBGFS2018/sol/w10/lbg_fs_2018_w10_sol09.pdf)"

# w12 sol/w11/lbg_fs_2018_w11_sol10.pdf
vecSolTopic[which(Date == format(as.Date("2018/12/07"), "%d.%m"))] <- "[Inbreeding and Henderson's Rule](https://charlotte-ngs.github.io/LBGFS2018/sol/w11/lbg_fs_2018_w11_sol10.pdf)"

# w13 sol/w12/lbg_fs_2018_w12_sol11.pdf
vecSolTopic[which(Date == format(as.Date("2018/12/14"), "%d.%m"))] <- "[Accuracy and Decomposition of PBV](https://charlotte-ngs.github.io/LBGFS2018/sol/w12/lbg_fs_2018_w12_sol11.pdf)"

# w14 
vecSolTopic[which(Date == format(as.Date("2018/12/21"), "%d.%m"))] <- "[Final Exam 2017](https://charlotte-ngs.github.io/LBGFS2018/sol/w13/lbg_hs2017_exmsol.pdf), [Multivariate BLUP and Variance Components](https://charlotte-ngs.github.io/LBGFS2018/sol/w13/lbg_fs_2018_w13_sol12.pdf)"


### # construct the table
dfSolTable <- data.frame(Woche, Date, vecSolTopic, stringsAsFactors = FALSE)
### # generate table output
colnames(dfSolTable) <- vecTableHeaders
knitr::kable(dfSolTable)
```


### More Material
```{r FurtherMatTable, echo=FALSE, results='asis'}
### # define vector for subjects
vecAddMatTopic = vector(mode = "character", length = nNrSemesterWeeks)
### # w1
vecAddMatTopic[which(Date == format(as.Date("2018/09/21"), "%d.%m"))] <- "[Introduction to Linear Algebra and to R/RStudio](https://charlotte-ngs.github.io/LBGFS2018/cn/91-app-intro.pdf)"

# w4 w04/lbg_fs_2018_w04_ex3.pdf
vecAddMatTopic[which(Date == format(as.Date("2018/10/12"), "%d.%m"))] <- "[Iris Dataset](https://charlotte-ngs.github.io/LBGFS2018/ex/w04/iris.csv)"

# w5 misc/20181019_special_matrices/20181019_special_matrices.pdf
vecAddMatTopic[which(Date == format(as.Date("2018/10/19"), "%d.%m"))] <- "[Special Matrices](https://charlotte-ngs.github.io/LBGFS2018/misc/20181019_special_matrices/20181019_special_matrices.pdf)"

# w8 misc/20181109_correction_section621/20181109_correction_section621.pdf
vecAddMatTopic[which(Date == format(as.Date("2018/11/09"), "%d.%m"))] <- "[Correction for Section 6.2.1](https://charlotte-ngs.github.io/LBGFS2018/misc/20181109_correction_section621/20181109_correction_section621.pdf)"

# w09 misc/20181116_correct_mendelian_variance/20181116_correct_mendelian_sampling.pdf
vecAddMatTopic[which(Date == format(as.Date("2018/11/16"), "%d.%m"))] <- "[Correction for Mendelian Sampling Term](https://charlotte-ngs.github.io/LBGFS2018/misc/20181116_correct_mendelian_variance/20181116_correct_mendelian_sampling.pdf)"

# w10 misc/20181116_verify_ldl_of_a/20181116_verify_ldl_of_a.nb.nb.nb.html
vecAddMatTopic[which(Date == format(as.Date("2018/11/23"), "%d.%m"))] <- "[Verify LDL-decomposition of A](https://charlotte-ngs.github.io/LBGFS2018/misc/20181116_verify_ldl_of_a/20181116_verify_ldl_of_a.nb.nb.nb.html)"

# w11 sol/w05/lbg_fs_2018_w05_sol04_corrected.pdf
vecAddMatTopic[which(Date == format(as.Date("2018/11/30"), "%d.%m"))] <- "[Correction in Solution to Exercise 4](https://charlotte-ngs.github.io/LBGFS2018/sol/w05/lbg_fs_2018_w05_sol04_corrected.pdf)"

# exam solutions
vecAddMatTopic[which(Date == format(as.Date("2018/12/21"), "%d.%m"))] <-"[Solutions to Final Exam](https://charlotte-ngs.github.io/LBGFS2018/exam/lbg_fs2018_exmsol.pdf)" 

# post exam solutions
Date <- c(Date, format(as.Date("2019/01/14"), "%d.%m"))
Woche <- c(Woche,NA)

vecAddMatTopic[which(Date == format(as.Date("2019/01/14"), "%d.%m"))] <-"[Solutions to Post-Final Exam](https://charlotte-ngs.github.io/LBGFS2018/exam/lbg_fs2018_postexmsol.pdf)" 

### # construct the table
dfAddMatTable <- data.frame(Woche, Date, vecAddMatTopic, stringsAsFactors = FALSE)

### # generate table output
colnames(dfAddMatTable) <- vecTableHeaders
knitr::kable(dfAddMatTable)
```

```{r LatestChange, echo=FALSE, results='asis'}
cat(rmddochelper::get_latest_change(), "\n")
```


